\documentclass{beamer}
%\documentclass[xcolor=pst,dvips,epic,eepic]{beamer}

  \usetheme{Warsaw}

%\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
%\usepackage[pst]{xcolor}
%\usepackage{xxcolor}
\usepackage{mathlist}

\usepackage[frenchb]{babel}
\usepackage[latin1]{inputenc}
\usepackage{amsmath,amssymb}
%\usepackage{colortbl}
%\usepackage[english]{babel}
\usepackage{pstricks, pst-tree}

\usepackage{times}

%\usepackage{pst-tree}

\usepackage{ulem}

\usepackage{listings}
%\topmargin=-0.6in
\lstloadlanguages{C++}
\lstset{language=C++}
%%}

\setbeamercovered{dynamic}

\def\bxi{\boldsymbol\xi}
\def\bx{\boldsymbol{x}}
\def\bu{\boldsymbol{u}}
\def\bz{\boldsymbol{z}}
\def\bomega{\boldsymbol\omega}

\title[Multi-stage problems]{Stochastic programming\\Multi-stage problems}

\author[Fabian Bastin]{Fabian Bastin\\\url{fabian.bastin@cirrelt.ca}\\Université de Montréal -- CIRRELT}

\date{}

\setbeamertemplate{footline}[frame number]

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Motivation: example}

\begin{itemize}
\item
Assume that we invest in the market.
We would to reach a nominal goal $G$ after $Y$ years.
\item
The portfolio is rebalanced each $v$ years, so there are $T = Y/v$ times where we have to decide what are the investments to do.\\
There are $T$ stages in our problem.
\item
We face a space of $N$ investment decisions (i.e. we have $N$ possible actions); we have a set $\mathcal{T} = \lbrace 1, 2,\ldots, T \rbrace$ of investment periods.
\item
If we go over the target $G$, we will have an interest rate $q$ on the surplus.
\item
If we do not meet the target, we have to borrow the difference at an interest $r$.
\item
The initial wealth is $b$.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Variables}

\begin{itemize}
\item
$x_{it}$, $i \in N$, $t \in \mathcal{T}$: amount of money to invest in action $i$ during the period $t$;
\item
$y$: excess amount at the end of the horizon;
\item
$w$: lack of money at the end of the horizon;
\item
In the current framework, we aim to take the corrective actions in order to reach the target of $G$ dollars.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Deterministic formulation}

\begin{align*}
\max\ qy & - rw\\
\mbox{s.t. } \sum_{i \in N} x_{i1} & = b,\\
\sum_{i \in N} \omega_{it}x_{i,t-1} & = \sum_{i \in N} x_{it}, \quad
\forall t \in \mathcal{T} \backslash 1, \\
\sum_{i \in N} \omega_{iT}x_{iT} - y + w & = G,\\
x_{it} & \geq 0, \quad \forall i \in N,\ t \in \mathcal{T}, \\
y, w & \geq 0.
\end{align*}

\end{frame}

\begin{frame}
\frametitle{Random returns}

\begin{itemize}
\item
The investment returns can be viewed as random variables, as we do not know their realizations in advance.
\item
Imagine that for each action, at each time $t$, the are $R$ possible realizations.
\item
The scenarios are made of all the possible sequences of realizations.
\end{itemize}

Example: assume $R = 4$ and $T = 3$. Scenarios:
\begin{small}
\begin{center}
\begin{tabular}{ccc}
\hline
$t = 1$ & $t = 2$ & $t = 3$ \\
\hline
1 & 1 & 1 \\
1 & 1 & 2 \\
1 & 1 & 3 \\
1 & 1 & 4 \\
1 & 2 & 1 \\
\vdots & \vdots & \vdots \\
4 & 4 & 4 \\
\end{tabular}
\end{center}
\end{small}

\end{frame}

\begin{frame}
\frametitle{Stochastic program formulation}

\begin{itemize}
\item
Let $x_{its}$, $t \in \mathcal{T}$, $s \in S$ be the amount to invest in action $i$ during the period $t$ in scenario $s$;
\item
$y_s$: excess amount at the end of the horizon for scenario $s$;
\item
$w_s$: lack of money at the end of the horizon for scenario $s$.
\item
Let $\omega_{it}$ be the random return; it can be seen as function of the scenario $s$.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Stochastic version}

\begin{align*}
\max\ \sum_{s=1}^S p_s(qy_{\red s} & - rw_{\red s})\\
\mbox{s.t. } \sum_{i \in N} x_{i1} & = b,\\
\sum_{i \in N} \omega_{it{\red s}}x_{i,t-1,{\red s}} & = \sum_{i \in
  N} x_{it{\red s}}, \quad \forall t \in \mathcal{T} \backslash 1,
{\red \forall s \in S} \\
\sum_{i \in N} \omega_{iT{\red s}}x_{iT} - y_{\red s} + w_{\red s} & = G,
\quad {\red \forall s \in S}, \\
x_{it{\red s}} & \geq 0, \quad \forall i \in N,\ t \in \mathcal{T},
{\red \forall s \in S},\\
y_{\red s}, w_{\red s} & \geq 0 \quad {\red \forall s \in S}.
\end{align*}

\end{frame}

\begin{frame}
\frametitle{Scenario tree}

Conceptually, the sequence of events can be arranged in a tree.
There is one leaf per scenario.
\begin{center}
%\begin{footnotesize}
\scalebox{0.45}{
\pstree[treemode=D]{\TC}{
 \pstree{\TC}
  {\TC~[tnpos=b]{\quad\bf{\large {\blue $1$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $2$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $3$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $4$}}}}
 \pstree{\TC}
  {\TC~[tnpos=b]{\quad\bf{\large {\blue $5$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $6$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $7$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $8$}}}}
 \pstree{\TC}
  {\TC~[tnpos=b]{\quad\bf{\large {\blue $9$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $10$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $11$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $12$}}}}
 \pstree{\TC}
  {\TC~[tnpos=b]{\quad\bf{\large {\blue $13$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $14$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $15$}}}
   \TC~[tnpos=b]{\quad\bf{\large {\blue $16$}}}}
}
}
%\end{footnotesize}
\end{center}

The scenarios can have different probabilities, the variables can be correlated,\ldots

\end{frame}

\begin{frame}
\frametitle{Non-anticipativity}

\begin{itemize}
\item
The conversion between the deterministic model and the stochastic version is easy\ldots
\item
but the built model is wrong!!!
\item
We have to enforce the {\red non-anticipativity}:
\end{itemize}

Let $S_s^t$ be the set of scenarios that are identical to scenario $s$ at time $t$.
We have to guarantee that
\[
x_{its} = x_{its'},\ \forall i \in N,\ \forall t \in T,\ \forall s \in
S,\ \forall s' \in S_s^t.
\]

\end{frame}

\begin{frame}
\frametitle{Non-anticipativity (cont'd)}

Equivalently,
\[
\left( \sum_{s' \in S_s^t} p'_s \right) x_{its} = \sum_{s' \in S_s^t} p'_s x_{its'} \quad \forall\ i,t,s,
\]
or
\begin{align*}
 x_{its}
&= \frac{\sum_{s' \in S_s^t} p'_s x_{its'}}{\sum_{s' \in S_s^t} p'_s} \quad \forall\ i,t,s \\
&= E[ x_{its'} | s' \in S_s^t ] \quad \forall\ i,t,s.
\end{align*}

The amount invested in scenario $s$, multiplied by the probability to reach a scenario equivalent to $s$, has to be equal to the expected amount that would be invested in any scenario equivalent to $s$.

There are many formulations of these conditions!

\end{frame}

\begin{frame}
\frametitle{Back to the scenario tree}

\begin{itemize}
\item
We have to enforce nonanticipativity as we have created copies of the variables for each scenario, at each period.
Graphically:
\end{itemize}
\begin{minipage}{0.45\textwidth}
\begin{center}
%\begin{footnotesize}
\hspace*{1cm} \scalebox{0.55}{
\pstree[treemode=R]{\TC}{
 \pstree{\TC}
  {\TC~[tnpos=r]{\quad\bf scenario 1}
   \TC~[tnpos=r]{\quad\bf scenario 2}
   \TC~[tnpos=r]{\quad\bf scenario 3}}
 \pstree{\TC}
  {\TC~[tnpos=r]{\quad\bf scenario 4}}
 \pstree{\TC}
  {\TC~[tnpos=r]{\quad\bf scenario 5}
   \TC~[tnpos=r]{\quad\bf scenario 6}}
}
}
%\end{footnotesize}
\end{center}
\end{minipage}
\hfill
\begin{minipage}{0.45\textwidth}
%\setlength{\unitlength}{0.69mm}
\begin{center}
\scalebox{0.7}{
\begin{pspicture}(4,0)(3,5)
\pscircle[linewidth=1pt](0,5){0.2}
\psline(0.21,5)(1.79,5)
\pscircle[linewidth=1pt](2,5){0.2}
\psline(2.21,5)(3.79,5)
\pscircle[linewidth=1pt](4,5){0.2}
\psline(-0.04,4.76)(-0.04,4.24)
\psline(0.04,4.76)(0.04,4.24)
\psline(1.96,4.76)(1.96,4.24)
\psline(2.04,4.76)(2.04,4.24)
\pscircle[linewidth=1pt](0,4){0.2}
\psline(0.21,4)(1.79,4)
\pscircle[linewidth=1pt](2,4){0.2}
\psline(2.21,4)(3.79,4)
\pscircle[linewidth=1pt](4,4){0.2}
\psline(-0.04,3.76)(-0.04,3.24)
\psline(0.04,3.76)(0.04,3.24)
\psline(1.96,3.76)(1.96,3.24)
\psline(2.04,3.76)(2.04,3.24)
\pscircle[linewidth=1pt](0,3){0.2}
\psline(0.21,3)(1.79,3)
\pscircle[linewidth=1pt](2,3){0.2}
\psline(2.21,3)(3.79,3)
\pscircle[linewidth=1pt](4,3){0.2}
\psline(-0.04,2.76)(-0.04,2.24)
\psline(0.04,2.76)(0.04,2.24)
\pscircle[linewidth=1pt](0,2){0.2}
\psline(0.21,2)(1.79,2)
\pscircle[linewidth=1pt](2,2){0.2}
\psline(2.21,2)(3.79,2)
\pscircle[linewidth=1pt](4,2){0.2}
\psline(-0.04,1.76)(-0.04,1.24)
\psline(0.04,1.76)(0.04,1.24)
\pscircle[linewidth=1pt](0,1){0.2}
\psline(0.21,1)(1.79,1)
\pscircle[linewidth=1pt](2,1){0.2}
\psline(2.21,1)(3.79,1)
\pscircle[linewidth=1pt](4,1){0.2}
\psline(-0.04,0.76)(-0.04,0.24)
\psline(0.04,0.76)(0.04,0.24)
\psline(1.96,0.76)(1.96,0.24)
\psline(2.04,0.76)(2.04,0.24)
\pscircle[linewidth=1pt](0,0){0.2}
\psline(0.21,0)(1.79,0)
\pscircle[linewidth=1pt](2.0,0){0.2}
\psline(2.21,0)(3.79,0)
\pscircle[linewidth=1pt](4.0,0){0.2}
\end{pspicture}}
\end{center}
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Nodal approach}

\begin{itemize}
\item
We can also create a vector of variables for each node in the tree.
\item
This vector corresponds to what should be our decision given the known realizations of the random variables upon this stage (i.e. at this node).
\item
Index the nodes by $l = 1,2,\ldots,\mathcal{L}$.
\item
We have to know the ancestor of each node.
Let $A(l)$ the ancestor of node $l \in \mathcal{L}$ in the scenarios tree.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Reformulation}

\begin{align*}
\max\ \sum_{s \in S} p_s (qy_{s} & - rw_{s})\\
\mbox{s.t. } \sum_{i \in N} x_{i1} & = b,\\
\sum_{i \in N} \omega_{il}x_{i,A(l)} & = \sum_{i \in N} x_{il}, \quad \forall l \in \mathcal{L} \backslash 1, \\
\sum_{i \in N} w_{iA(s)}x_{iA(s)} - y_{s} + w_{s} & = G,
\quad {\forall s \in S}, \\
x_{il} & \geq 0, \quad \forall i \in N,\ \forall l \in \mathcal{L},\\
y_{s}, w_{s} & \geq 0 \quad {\forall s \in S}.
\end{align*}

\end{frame}

\begin{frame}
\frametitle{Multi-period production planification}

\begin{itemize}
\item
A firm produces various different goods.
\item
Ressources (for instances manual and machine work hours) are required.
We assume that these requirements are known for each product.
\item
The demand must be met at the end of each period, but this demand is not perfectly known.
\item
A too large or too small inventory induces costs (storage, additional purchases to satisfy the demand,\ldots).
\item
It is possible to add machine and work hours, within predefined limits.
\item
There is a hiring and firing cost related to changes in the workforce.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Decision problem}

\begin{itemize}
\item
We have to decide now on
\begin{itemize}
\item
the quantity of each product to manufacture during each period;
\item
the additional capacity to use during each period;
\item
to hire or to fire staff during each period;
\end{itemize}
\item
A random demand occurs. Conceptually, this happens at each period.
\item
After having observed the random demands, we can decide on how to stock the products in the inventory, or to buy from an external source.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Definitions}

\begin{enumerate}
\item
{\red Sets}:
\begin{itemize}
\item
$T$: number of periods, also seen as a set;
\item
$N$: products;
\item
$M$: ressources.
\end{itemize}
\item
{\red Variables}:
\begin{itemize}
\item
$x_{jt}$: production of product $j \in N$ during the period $t \in T$;
\item
$u_{it}$: additional amount of ressource $i \in M$ to buy during the period  $t \in T$;
\item
$z_{t-1,t}^+$, $z_{t-1,t}^-$: planned augmentation or reduction of the labor force;
\item
$y_{j,t}^+$, $y_{j,t}^-$: surplus, deficit of product $j \in N$ at the end of the period $t \in T$.
\end{itemize}
\end{enumerate}

\end{frame}

\begin{frame}
\frametitle{Parameters}

\begin{itemize}
\item
All the above variables have associated costs $(\alpha,\,\beta,\,\gamma,\,\delta)$.
\item
$\omega_{jt}$: demand of product $j \in N$ during the period $t
\in T$.
\item
$U_{it}$: upper bound on $u_{it}$.
\item
$a_{ij}$: amount of ressource $i \in M$ required to produce one unit of the good $j \in N$.
\item
$b_{it}$: amount of ressource $i \in M$ available at time $t \in T$.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Model}

\begin{multline*}
\min \sum_{j \in N} \sum_{t \in T} \alpha_{jt} x_{jt} + \sum_{i \in M}
\sum_{t \in T} \beta_{it} u_{it} + \sum_{t \in T\backslash 1} \left(
  \gamma^+_{t-1,t} z^+_{t-1,t} + \gamma^-_{t-1,t} z^-_{t-1,t} \right) \\
+ E_{\bomega} \left[ \min \left( \sum_{j \in N} \sum_{t \in T}
    ( \delta_{jt}^+y_{jt}^+ + \delta_{jt}^-y_{jt}^- ) \right) \right],
\end{multline*}
such that
\begin{align*}
\sum_{j \in N} a_{ij} x_{jt} & \leq b_{it}+u_{it} \quad \forall i \in
M,\ \forall t \in T,\\
u_{it} & \leq U_{it} \qquad \forall i \in M,\ \forall t \in T,\\
z^+_{t-1,t} - z^-_{t-1,t} &= \sum_{j \in N} a_{jt}(x_{jt} - x_{j,
  t-1}) \quad \forall t \in T\backslash 1, \\
x_{jt} + y_{j,t-1}^+ - y_{jt}^+ + y_{jt}^- &= \omega_{jt} \qquad
\forall j \in N,\ \forall t \in T\ \mbox{(demand satisfaction)},\\
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Extended form of the recourse model}

\begin{multline*}
\min \sum_{j \in N} \sum_{t \in T} \alpha_{jt} \bx_{jt} + \sum_{i \in M}
\sum_{t \in T} \beta_{it} \bu_{it} + \sum_{t \in T\backslash 1} \left(
  \gamma^+_{t-1,t} \bz^+_{t-1,t} + \gamma^-_{t-1,t} \bz^-_{t-1,t} \right) \\
+ {\red \sum_{s \in S}}\sum_{j \in N} \sum_{t \in T} p_s
    ( \delta_{jt}^+y_{jt{\red s}}^+ + \delta_{jt}^-y_{jt{\red s}}^- ),
\end{multline*}
such that
\begin{align*}
\sum_{j \in N} a_{ij} \bx_{jt} & \leq b_{it}+\bu_{it} \quad \forall i \in
M,\ \forall t \in T,\\
\bu_{it} & \leq U_{it} \qquad \forall i \in M,\ \forall t \in T,\\
\bz^+_{t-1,t} - \bz^-_{t-1,t} &= \sum_{j \in N} a_{jt}(\bx_{jt} - \bx_{j,
  t-1}) \quad \forall t \in T\backslash 1, \\
\bx_{jt} + y_{j,t-1,{\red s}}^+ - y_{jt{\red s}}^+ + y_{jt{\red s}}^- &= \omega_{jt{\red s}} \qquad
\forall j \in N,\ \forall t \in T,\\
\mbox{nonanticipativity }& \mbox{constraints}.
\end{align*}
\end{frame}

\end{document}
