\documentclass{beamer}
%\documentclass[xcolor=pst,dvips,epic,eepic]{beamer}

\usepackage[utf8]{inputenc}
\usetheme{Singapore}
\usepackage{xcolor}
\setbeamertemplate{footline}[frame number]

%\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
%\usepackage[pst]{xcolor}
%\usepackage{xxcolor}
\usepackage{mathlist}

\usepackage[utf8]{inputenc}
\usepackage[frenchb]{babel}
\usepackage{amsmath,amssymb}
%\usepackage{colortbl}
%\usepackage[english]{babel}
\usepackage{pstricks}
%\usepackage{pictex}

\usepackage{times}

%\usepackage{pst-tree}

\usepackage{ulem}

\usepackage{listings}
%\topmargin=-0.6in
\lstloadlanguages{C++}
\lstset{language=C++}
%%}

\setbeamercovered{dynamic}

\def\bxi{\boldsymbol\xi}
\def\bomega{\boldsymbol\omega}

\title[Plan]{Introduction to Stochastic optimization}

\author{Fabian Bastin}
\author[Fabian Bastin]{Fabian Bastin \\ \url{fabian.bastin@cirrelt.ca} \\ Université de Montréal -- CIRRELT -- IVADO --Fin-ML}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Goals}

%\section{Objectifs}

{\bf Main goal}: introduction to stochastic optimization, that aims to optimize (in the mathematical way) while taking the uncertainty into account.

\mbox{}

{\bf Plan}
\begin{itemize}
\item
uncertainty and optimization;
\item
stochastic programming with recourse;
\item
two-stage linear stochastic programming, compact and extended formulations, L-Shaped method;
\item
multistage problems;
\item
convex and non convex problems
\item
sampling methods and consistency analysis;
\item
stochastic gradient descent -- stochastic approximation.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Organization}

{\bf Prerequise}
\begin{itemize}
\item
good mathematical and statistical background;
\item
it is advised to have some background in optimization;
\item
basic knowledge in programming.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{References}

{\bf Main reference}
\begin{itemize}
\item
John R. Birge et François Louveaux, {\sl Introduction to Stochastic
  Programming}, Springer-Verlag, 1997.
\end{itemize}

{\bf Suggested books}
\begin{itemize}
\item
Peter Kall et Stein W. Wallace, {\sl Stochastic Programming}, John
  Wiley \& Sons, 1994.
\item
Andrzej Ruszczynski et Alexander Shapiro, {\sl Stochastic
  Programming}, Elsevier, 2003.
\end{itemize}

{\bf Software}
\begin{itemize}
\item
Julia (\url{http://www.julialang.org}),
\end{itemize}

{\bf Class material}
\begin{itemize}
	\item
	\url{https://github.com/fbastin/SP_Introduction}
\end{itemize}


\end{frame}

\begin{frame}
\frametitle{Evaluation}

The evaluation modalities will be discussed in the class. The main elements are
\begin{itemize}
\item
2 homeworks (20\% each);
\item
1 project (30\%)
\item
oral presentation of a scientific paper (30\%)
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Objectives}
 
\underline{Introduction} to stochastic optimization, i.e.
\begin{itemize}
\item
get familiar with the terminology, the possibilities and main limitations of stochastic programming models;
\item
learn how to formulate analytical models incorporating uncertainty as stochastic programs;
\item
learn the basic theory;
\item
learn the main algorithms used to solved the associated problems.
\end{itemize}
\end{frame}


\end{document}
